Towards Practical First-Order Model Counting

### 2024-02-14 doc/paper.tex
Ditto
Experimental setup for ForcLift (just for a linearly-increasing sequence of domains for now)

\todo[inline,caption={}]{
  \begin{itemize}
    \item mention/describe the arithmetic simplification code
    \item (in Fig. 1) Crane2 doesn't use Crane, it uses itself
    \item describe which compilation rules are currently greedy or non-greedy
  \end{itemize}
}

\paragraph{Changes to the compilation process.}
\begin{itemize}
  \item We make the FCGs satisfy the following property. For every domain
        $\Delta$, for every directed path $P$ without Ref nodes, the number of
        atom counting or (generalised) domain recursion or independent partial
        grounding operations on $\Delta$ on $P$ is at most one.
  \item when checking whether recursion is possible between a source and a
        target formula, the target formula (maybe the source formula as well?)
        must not include a domain as well as its subdomain. This might be
        equivalent to the formula being `unshattered'. Such formulas could be
        seen as transitory, and allowing such recursive calls leads to FCGs that
        I cannot make sense of.
  \item During smoothing, when unit clauses (a.k.a.\ variables) are propagated
        in the opposite direction of the FCG arcs (i.e., `upwards'), when
        visiting a \texttt{Ref} node, these clauses are translated using the
        domain map of the \texttt{Ref} node. For example, we might have a
        function $g$ that depends on $m$ (i.e., domain $a^{\top}$), which is
        called as $g(n)$ (i.e., on domain $a$). When smoothing that function
        call, we want to translate all mentions of $a^{\top}$ to $a$.
\end{itemize}

\paragraph{Proof of correctness.}
\begin{enumerate}
  \item Define what it means for a formula to be \emph{consistent}.
        \begin{itemize}
          \item The formula (including the mapping of constants to their
                domains) does not mention a domain together with its subdomain
                (i.e., all mentioned domains are pairwise disjoint).
                Specifically, we can't have a subdomain together with its parent
                domain or a subdomain together with a constant that belongs to
                its parent domain.
          \item The target of \texttt{Ref} must also be a consistent formula.
          \item The initial formula is consistent by definition.
        \end{itemize}
  \item Aim to show that Crane works correctly on all consistent formulas.
  \item Prove that, whenever some compilation rule makes the formula $\phi$
        inconsistent, greedy compilation rules will make $\phi$ consistent again
        before $\phi$ encounters non-greedy compilation rules that depend on
        $\phi$ being consistent.
        \begin{itemize}
          \item In particular, prove that shattering and unit propagation will
                always eliminate the parent domain of the domains introduced by
                atom counting.
          \item Constraint removal only applies when it can eliminate the parent
                domain entirely.
          \item Prove this for a particular assumption about which compilation
                rules are set to be greedy. For the greedy algorithm, although
                all rules are applied in a greedy manner, we still assume that
                greedy rules will be applied before non-greedy rules.
        \end{itemize}
  \end{enumerate}

@@

\section{The Main Algorithm: Finding the Definitions of the Base Cases}

@@

>>>>

\todo[inline,caption={}]{
  \begin{itemize}
    \item mention/describe the arithmetic simplification code
    \item (in Fig. 1) Crane2 doesn't use Crane, it uses itself
    \item describe which compilation rules are currently greedy or non-greedy
    \item my agenda has a plan for proving completeness for $\mathbf{C}^{2}$
  \end{itemize}
}

\paragraph{Changes to the compilation process.}
\begin{itemize}
  \item We make the FCGs satisfy the following property. For every domain
        $\Delta$, for every directed path $P$ without Ref nodes, the number of
        atom counting or (generalised) domain recursion or independent partial
        grounding operations on $\Delta$ on $P$ is at most one.
  \item when checking whether recursion is possible between a source and a
        target formula, the target formula (maybe the source formula as well?)
        must not include a domain as well as its subdomain. This might be
        equivalent to the formula being `unshattered'. Such formulas could be
        seen as transitory, and allowing such recursive calls leads to FCGs that
        I cannot make sense of.
  \item During smoothing, when unit clauses (a.k.a.\ variables) are propagated
        in the opposite direction of the FCG arcs (i.e., `upwards'), when
        visiting a \texttt{Ref} node, these clauses are translated using the
        domain map of the \texttt{Ref} node. For example, we might have a
        function $g$ that depends on $m$ (i.e., domain $a^{\top}$), which is
        called as $g(n)$ (i.e., on domain $a$). When smoothing that function
        call, we want to translate all mentions of $a^{\top}$ to $a$.
  \item Implemented an if-then-else arithmetic operation as the interpretation
        of the new domain recursion node (i.e., there is one branch for when the
        domain is empty and another branch for any other domain size).
  \item Smoothing for domain recursion now works the same way as smoothing for
        disjunction: both sides must have the same atoms. Almost all of
        \cref{sec:simplifying} is now outdated because smoothing makes the
        ``tautology insertion'' unnecessary.
  \item \Cref{sec:main,sec:identifying} might also be outdated: there is
        probably no need to search for base cases given the new version of
        domain recursion.
\end{itemize}

\paragraph{Proof of correctness.}
\begin{enumerate}
  \item Define what it means for a formula to be \emph{consistent}.
        \begin{itemize}
          \item The formula (including the mapping of constants to their
                domains) does not mention a domain together with its subdomain
                (i.e., all mentioned domains are pairwise disjoint).
                Specifically, we can't have a subdomain together with its parent
                domain or a subdomain together with a constant that belongs to
                its parent domain.
          \item The target of \texttt{Ref} must also be a consistent formula.
          \item The initial formula is consistent by definition.
        \end{itemize}
  \item Aim to show that Crane works correctly on all consistent formulas.
  \item Prove that, whenever some compilation rule makes the formula $\phi$
        inconsistent, greedy compilation rules will make $\phi$ consistent again
        before $\phi$ encounters non-greedy compilation rules that depend on
        $\phi$ being consistent.
        \begin{itemize}
          \item In particular, prove that shattering and unit propagation will
                always eliminate the parent domain of the domains introduced by
                atom counting.
          \item Constraint removal only applies when it can eliminate the parent
                domain entirely.
          \item Prove this for a particular assumption about which compilation
                rules are set to be greedy. For the greedy algorithm, although
                all rules are applied in a greedy manner, we still assume that
                greedy rules will be applied before non-greedy rules.
        \end{itemize}
  \item In particular, independent partial groundings and atom counting are the
        two classical rules that require the formula to be shattered.
  \end{enumerate}

@@

\section{The Main Algorithm: Finding the Definitions of the Base
  Cases}\label{sec:main}

@@

\paragraph{Remarks on comparing the model counts of the formulas with their
  corresponding OEIS sequences.}
Disregarding differences where the OEIS sequence starts with one or two numbers
that are not included in model counts (which is somewhat justified since we
start with domains of size one rather than zero), two differences remain:
\begin{itemize}
  \item on A62971, the OEIS sequence starts as $1, 2, 16,\dots$, but the model
        counts start as $1, 16,\dots$ (with the rest exactly the same);
  \item similarly, on A62971, the OEIS sequence starts as $32, 290,\dots$, but
        the model counts start as $2, 32, 290,\dots$ (with the rest exactly the
        same).
\end{itemize}
